---
title: "Exploratory analysis"
output:
  pdf_document: default
  html_notebook: default
---

## Attach packages
```{r, message = FALSE}
library(tidyverse)
library(caret)
library(psych)
library(BioStatR)
library(car)
library(lattice)
```

## Load the datafile

```{r}
daily_data <- readRDS("daily_data.rds")
summary(daily_data)

```

## Examine distributions

```{r}
numeric <- daily_data[,3:10]
multi.hist(numeric)
```

The target variable (meter_reading) is heavily right-skewed. cloud_coverage and precip_depth_1_hr are also right-skewed. 

## Examine correlations
```{r}
round(cor(numeric), 2)
```
None of the weather variables are strongly linearly correlated with meter-reading. A fact that we will see in the following plots.


## Examine scatterplots
```{r}
daily_data$meter <- as.factor(daily_data$meter)
air_temp <- ggplot(daily_data, aes(x = air_temperature, y = log(meter_reading), colour = meter)) + geom_point()+geom_smooth(method= lm) 
air_temp

```

There are outliers but for the most part there is not much of a linear relationship between air_temperature and meter_reading. meter type 3 does show a stronger linear relationship than the other two meter types.


```{r}
cloud_cover <- ggplot(daily_data, aes(x = cloud_coverage, y = log(meter_reading), colour = meter)) + geom_point()+geom_smooth(method= lm) 
cloud_cover

```
cloud_coverage does not have a clear linear relationship with meter reading except for meter 3.

```{r}
dew_temp <- ggplot(daily_data, aes(x = dew_temperature, y = log(meter_reading), colour = meter)) + geom_point()+geom_smooth(method= lm)
dew_temp

```
We see the same trend in dew temperature as the earlier two temperature variables. There is not much of a linear relationship, except for meter 3.

```{r}
precip_depth <- ggplot(daily_data, aes(x = precip_depth_1_hr, y = log(meter_reading), colour = meter)) + geom_point()+geom_smooth(method= lm)
precip_depth
```
Precip_depth_1_hr does not have any relationship with meter reading. Even meter 3 is all over the place.

```{r}
sea_pressure <- ggplot(daily_data, aes(x = sea_level_pressure, y = log(meter_reading), colour = meter)) + geom_point()+geom_smooth(method= lm)
sea_pressure
```
sea_level pressure has no relationship with meter reading for all meter types.

```{r}
wind_direction <- ggplot(daily_data, aes(x = wind_direction, y = log(meter_reading), colour = meter)) + geom_point()+geom_smooth(method= lm)
wind_direction
```
No relationship between wind direction and meter reading. meter 3 is all over the place!

```{r}
wind_speed <- ggplot(daily_data, aes(x = wind_speed, y = log(meter_reading), colour = meter)) + geom_point()+geom_smooth(method= lm)
wind_speed
```
No relationship between wind speed and meter reading.

## Look at the Building related variables

```{r}
building_vars <- readRDS("building_vars.rds")
summary(building_vars)
dim(building_vars)
```
```{r}
building_vars$primary_use <- as.factor(building_vars$primary_use)
summary(building_vars$primary_use)
meter_log <- log(building_vars$meter_reading)
use <- ggplot(building_vars, aes(x = primary_use, y = meter_log)) + geom_boxplot() +
   stat_summary(fun.y=mean, colour="darkred", geom="point", 
               shape=18, size=3)
use + theme(axis.text.x = element_text(angle = 40, hjust = 1))
```
Not much of a difference in means. I logged meter readings because they are so strongly right skewed. I could not see a trend if I did not log them. No relationship between use of building and meter reading.


Now, lets look at the area of the buildings. Again this variable is highly right-skewed. I'm going to log it.
```{r}
hist(building_vars$square_feet)
```
```{r}
area_log <- log(building_vars$square_feet)
hist(area_log)
```
looks much better! 

```{r}
plot(area_log, meter_log)
```
Nothing going on here! There is no relationship between area and meter readings. 


```{r}
building_vars$year_built <- as.factor(building_vars$year_built)
summary(building_vars$year_built)
meter_log <- log(building_vars$meter_reading)
year <- ggplot(building_vars, aes(x = year_built, y = meter_log)) + geom_boxplot() + stat_summary(fun.y=mean, colour="darkred", geom="point", shape=18, size=3)
year + theme(axis.text.x = element_text(size = 5, angle = 90, hjust = 1))
```
There is some variation but not much of a trend. It almost seems random. Surprisingly, the newer buildings are not using less energy!!

Now lets look at floor count!

```{r}
hist(building_vars$floor_count)
```
```{r}
table(building_vars$floor_count)
```
High number of 1 storey buildings. Lets cut this into a factor with 1 storey, 2 storey, 3 to 9 storey and more than 9 storey buildings. That looks like a natural grouping to me looking at the histogram. 

```{r}
cut_storey <- cut(building_vars$floor_count, breaks = c(1, 2, 3, 9, 26))

summary(cut_storey)
cut_storey <- as.factor(cut_storey)
plot(cut_storey, meter_log)

```
Not much of a difference in the medians!

Main conclusion: My EDA does not show any strong relationships between the predictors and the target in this dataset. 

