---
title: "Exploratory analysis"
output: html_notebook
---

## Attach packages
```{r, message = FALSE}
library(tidyverse)
library(caret)
library(psych)
library(BioStatR)
library(car)
library(lattice)
library(lubridate)
```

## Load the datafile

```{r}
daily_data <- readRDS("daily_data.rds")
summary(daily_data)
```

```{r}

```


## Examine distributions

```{r}
numeric <- daily_data[,3:13]
multi.hist(numeric)
```

The target variable (meter_reading) is heavily right-skewed. cloud_coverage and precip_depth_1_hr are also right-skewed. 

## Examine scatterplots
```{r}
air_temp <- ggplot(daily_data, aes(x = air_temperature, y = meter_reading, colour = meter)) + geom_point()+geom_smooth(method= lm)
air_temp

```

There are outliers but for the most part there is not much of a linear relationship between air_temperature and meter_reading. meter type 3 does show a stronger linear relationship than the other two meter types.


```{r}
cloud_cover <- ggplot(daily_data, aes(x = cloud_coverage, y = meter_reading, colour = meter)) + geom_point()+geom_smooth(method= lm)
cloud_cover

```
cloud_coverage does not have a clear linear relationship with meter reading except for meter 3.

```{r}
dew_temp <- ggplot(daily_data, aes(x = dew_temperature, y = meter_reading, colour = meter)) + geom_point()+geom_smooth(method= lm)
dew_temp

```
We see the same trend in dew temperature as the earlier two temperature variables. There is not much of a linear relationship, except for meter 3.

```{r}
precip_depth <- ggplot(daily_data, aes(x = precip_depth_1_hr, y = meter_reading, colour = meter)) + geom_point()+geom_smooth(method= lm)
precip_depth
```
Precip_depth_1_hr does not have any relationship with meter reading. Even meter 3 is all over the place.

```{r}
sea_pressure <- ggplot(daily_data, aes(x = sea_level_pressure, y = meter_reading, colour = meter)) + geom_point()+geom_smooth(method= lm)
sea_pressure
```
sea_level pressure has no relationship with meter reading for all meter types.

```{r}
wind_direction <- ggplot(daily_data, aes(x = wind_direction, y = meter_reading, colour = meter)) + geom_point()+geom_smooth(method= lm)
wind_direction
```
No relationship between wind direction and meter reading. meter 3 is all over the place!

```{r}
wind_speed <- ggplot(daily_data, aes(x = wind_speed, y = meter_reading, colour = meter)) + geom_point()+geom_smooth(method= lm)
wind_speed
```
No relationship between wind speed and meter reading.

